\pkg{ggenealogy}: An \proglang{R} Package for Visualizing Genealogical Data

### 2016-11-06 article.tex
binVector to bin

\begin{Code}
R> plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:3, pathEdgeCol = "red",
+    nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::
+    element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
\end{Code}

@@

In the code above, syntax from the \pkg{ggplot2} package was appended to the \code{plotPathOnAll()} function; this can be done for most \pkg{ggenealogy} functions. While the first three explicit parameters have been introduced earlier in this paper, the fourth parameter (\code{binVector}) requires some explanation. The motivation of the \code{plotPathOnAll()} function is to write node labels on a plot, with the center of each node label constricted on the horizontal axis to its quantitative variable of interest (in this case, year of release). As is the case for the plots before, the vertical axis has no meaning other than providing a plotting area in which to draw the node labels. Unfortunately, for large datasets, this motivation can be a difficult task because the text labels of the varieties can overlap if they are assigned a similar y coordinate, have a similar year (x coordinate), and have long text labels (width of x coordinate).

For each variety, the x coordinate (year) and width of the x coordinate (text label width) cannot be altered, as they provide useful information. However, for each variety, the y coordinate is arbitrary. Hence, in an attempt to mitigate text overlapping, the \code{plotPathOnAll()} function does not randomly assign the y coordinate. Instead, it allows users to partially control the y coordinates with a user-determined number of bins (\code{binVector}).

@@

In the code above, \code{binVector} was assigned a value of 3, and \code{pathEdgeCol} was assigned a value of ``red". Additionally, we specified a size of 2.5 for the non-path node test using the \code{nodeSize} parameter, and a size of 4 for the path node text using the \code{pathNodeSize} parameter. There are several other parameters in the \code{plotPathOnAll()} function, which can be read in more detail using the help command.

@@

\begin{Code}
R> plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:6, pathEdgeCol =
+    "seagreen2", nodeSize = 1, pathNodeSize = 3)
\end{Code}

@@

We can achieve this using the below code. After trial and error, we use a \code{binVector} of size 200, and append \pkg{ggplot2} syntax to define suitable x-axis limits. The output of this process is illustrated in Figure \ref{fig:plotCBText}.

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200) +
+    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
+    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
+    (expand = c(.1, .2))
\end{CodeInput}
\end{CodeChunk}

@@

The text labels are so numerous in Figure \ref{fig:plotCBText} that simply trying different values for the input parameter \code{binVector} will not solve the text overlapping problem. Instead, one approach we can try is to reconstruct the plot using the same \pkg{ggenealogy} function \code{plotPathOnAll()}, only now specifying variables to render the size (2.5) and color (default of black) of the text for nodes that are on the path of interest to be more noticeable than the size (0.5) and color (dark gray) of the text for nodes that are not on the path of interest. Moreover, we can make the edges that are not on the path of interest to be represented in a less noticeable color (light gray) than the edges that are on the path of interest (default of dark green). The variable names and options for these aesthetics is further detailed in the help manual of the function. We provide one example code that alters the defaults of the text color and sizes of nodes and edges below, which results in Figure \ref{fig:plotCBNoText}.

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5,
+    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") +
+    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
+    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
+    (expand = c(.1, .2))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5,
+    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray",
+    animate = TRUE)
\end{CodeInput}
\end{CodeChunk}

>>>>

\begin{Code}
R> plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 1:3, pathEdgeCol = "red",
+    nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::
+    element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
\end{Code}

@@

In the code above, syntax from the \pkg{ggplot2} package was appended to the \code{plotPathOnAll()} function; this can be done for most \pkg{ggenealogy} functions. While the first three explicit parameters have been introduced earlier in this paper, the fourth parameter (\code{bin}) requires some explanation. The motivation of the \code{plotPathOnAll()} function is to write node labels on a plot, with the center of each node label constricted on the horizontal axis to its quantitative variable of interest (in this case, year of release). As is the case for the plots before, the vertical axis has no meaning other than providing a plotting area in which to draw the node labels. Unfortunately, for large datasets, this motivation can be a difficult task because the text labels of the varieties can overlap if they are assigned a similar y coordinate, have a similar year (x coordinate), and have long text labels (width of x coordinate).

For each variety, the x coordinate (year) and width of the x coordinate (text label width) cannot be altered, as they provide useful information. However, for each variety, the y coordinate is arbitrary. Hence, in an attempt to mitigate text overlapping, the \code{plotPathOnAll()} function does not randomly assign the y coordinate. Instead, it allows users to partially control the y coordinates with a user-determined number of bins (\code{bin}).

@@

In the code above, \code{bin} was assigned a value of 3, and \code{pathEdgeCol} was assigned a value of ``red". Additionally, we specified a size of 2.5 for the non-path node test using the \code{nodeSize} parameter, and a size of 4 for the path node text using the \code{pathNodeSize} parameter. There are several other parameters in the \code{plotPathOnAll()} function, which can be read in more detail using the help command.

@@

\begin{Code}
R> plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 1:6, pathEdgeCol =
+    "seagreen2", nodeSize = 1, pathNodeSize = 3)
\end{Code}

@@

We can achieve this using the below code. After trial and error, we use a \code{bin} of size 200, and append \pkg{ggplot2} syntax to define suitable x-axis limits. The output of this process is illustrated in Figure \ref{fig:plotCBText}.

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 1:200) +
+    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
+    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
+    (expand = c(.1, .2))
\end{CodeInput}
\end{CodeChunk}

@@

The text labels are so numerous in Figure \ref{fig:plotCBText} that simply trying different values for the input parameter \code{bin} will not solve the text overlapping problem. Instead, one approach we can try is to reconstruct the plot using the same \pkg{ggenealogy} function \code{plotPathOnAll()}, only now specifying variables to render the size (2.5) and color (default of black) of the text for nodes that are on the path of interest to be more noticeable than the size (0.5) and color (dark gray) of the text for nodes that are not on the path of interest. Moreover, we can make the edges that are not on the path of interest to be represented in a less noticeable color (light gray) than the edges that are on the path of interest (default of dark green). The variable names and options for these aesthetics is further detailed in the help manual of the function. We provide one example code that alters the defaults of the text color and sizes of nodes and edges below, which results in Figure \ref{fig:plotCBNoText}.

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 1:200, nodeSize = .5,
+    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") +
+    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
+    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
+    (expand = c(.1, .2))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 1:200, nodeSize = .5,
+    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray",
+    animate = TRUE)
\end{CodeInput}
\end{CodeChunk}
